---
title: "treeRchitecture_examples"
output: html_notebook
---

This notebook gives examples of operations on cylinder model of trees (QSMs)
It relies on the QSM format from Pasi Raumonen's TreeQSM
It is essentially a few summaries and plots from Allie Shenkin's treestruct 
+ a few functions from treeRchitecture_functions.R. I'm adding to this.

This section just sets up the packages
```{r}
#devtools::install_github("ashenkin/treestruct",force=TRUE)
#detach("package:treestruct", unload = TRUE)

library(treestruct)
library(rmatio)
library(testit)
library(R.matlab)
library(visNetwork)
library(tidyverse)
```

Load in the data and convert to treestructs following ashenkin github example
```{r}
TLS="C:/Users/tobyd/Documents/Data/TLS_Data/Malaysia/Sep09/test_QSMs/"

TLS="C:/Users/tobyd/Documents/Data/TLS_Data/QSMs/Global_QSMs/test_qsm_txt/"

my_trees = import_treestructs_from_dir(qsm_path = TLS, qsmver = "UCL", recursive = T)
my_qsms = TreeStructs(dataset = "my_trees", treestructs = my_trees)
my_qsms = run_all(my_qsms)  # run all architectural analyses
my_qsms = parse_id(my_qsms, treetag_regex = ".*(?=\\.mat)", nobranchcode = T)  # add tree tag to df

# Add tree tag to the cyl data frame inside the list
for (i in seq(1,length(my_qsms$treestructs$treestruct))){
  my_qsms$treestructs$treestruct[[i]]$tag=my_qsms$treestructs$tag[i]
}

my_treestructs = getTreestructs(my_qsms)  # tree level summaries of current architectural variables

# You can safely ignore the warnings below
```

Add 'treeRchitecture_functions'
```{r}
treeRchitecture="C:/Users/tobyd/OneDrive - University Of Cambridge/R/treeRchitecture/"
source(paste0(treeRchitecture,'treeRchitecture_functions.R'))
```


Below I calculate verticlaly binned summaries of the tree structure
I'm using the treestructs (rather than cyl_summs) so there are more cylinders per bin
It should be generalisable to cyl_summs once all the same variables are included there

```{r}

my_qsms$treestructs$treestruct %<>% lapply(add_centres_and_height_bin,bin_size=5) 
# bin_size = vertical size of the bins in m


rad_per_height_bin=bind_rows(my_qsms$treestructs$treestruct) %>% 
  group_by(tag,height_of_bin_m) %>% 
  summarize(max_rad=max(rad),median_rad=median(rad)) # Maximum radius cylinder per height bin

ggplot(rad_per_height_bin,aes(max_rad,height_of_bin_m,color=tag))+geom_line()

```


Calculate branching angles
```{r}
#tree=my_qsms$treestructs$treestruct[[1]] # This selects one tree
#tree=add_angles(tree) # add angles to this tree
# This function adds the angle each cylinder makes with it's parent

my_qsms$treestructs$treestruct %<>% lapply(add_angles) # add angles to all trees in the list
# this produces some NaN's where there is no parent cylinder, don't worry about that

# Plot the distributions of branching angles for the trees
bind_rows(my_qsms$treestructs$treestruct) %>%
  filter(order_in_branch==1) %>%  # use only cyls after the bifurcation points to get the branching angles only.
  ggplot(aes(x=angle_from_parent_z,color=tag))+stat_density(alpha = 1,size = 0.5,geom="line", position = "identity")

branching_angle_per_height_bin=bind_rows(my_qsms$treestructs$treestruct) %>% 
  group_by(tag,height_of_bin_m) %>% 
  filter(order_in_branch==1) %>% # select only the cyls after branching point to get branching angles
  summarize(mean_angle_z=mean(angle_from_parent_z,na.rm=TRUE),
            mean_angle_3d=mean(angle_from_parent_3d,na.rm=TRUE))  


ggplot(branching_angle_per_height_bin,aes(mean_angle_z,height_of_bin_m,color=tag))+
  geom_line()+geom_point()


```

Questions for Allie about treestruct
1. Can I import txt files?
2. Do cyl_summ trees have similar architectural properties to 'full' trees?
3. Why don't cyl_summs have the same variables as treestructs? (e.g. x_start)
4. Is there a function to get all daughter branches?
